---
title: "ECSC"
output: html_document
date: "2023-12-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries and data
```{r}

library(tidyverse)

df = read_csv("ECSC ECBI data.csv")



```


# Intensity scores
```{r}

#taking out useful columns

df_intensity = df %>% select(c(ecbi1_timestamp, Child_ID, ecbi1a, ecbi_2a, ecbi_3a, ecbi_4a, ecbi_5a, ecbi_6a, ecbi_7a, ecbi_8a, ecbi_9a, ecbi_10a, ecbi_11a, ecbi_12a, ecbi_13a, ecbi_14a, ecbi_15a, ecbi_16a, ecbi_17a, ecbi_18a, ecbi_19a, ecbi_20a, ecbi_21a, ecbi_22a, ecbi_22a_a, ecbi_24a, ecbi_25a, ecbi_26a, ecbi_27a, ecbi_28a, ecbi_29a, ecbi_30a, ecbi_31a, ecbi_32a, ecbi_33a, ecbi_34a, ecbi_35a, ecbi_36a))
```


# Problem Scores
```{r}

#taking out useful columns

df_problem = df %>% select(ecbi_1b, ecbi_2b, ecbi_3b, ecbi_4b, ecbi_5b, ecbi_6b, ecbi_7b, ecbi_8b, ecbi_9b, ecbi_10b, ecbi_11b, ecbi_12b, ecbi_13b, ecbi_14b, ecbi_15b, ecbi_16b,ecbi_17b, ecbi_18b, ecbi_19b, ecbi_20b, ecbi_21b, ecbi_22b, ecbi_22b_a, ecbi_24b, ecbi_25b, ecbi_26b, ecbi_27b, ecbi_28b, ecbi_29b, ecbi_30b, ecbi_31b, ecbi_32b, ecbi_33b, ecbi_34b, ecbi_35b, ecbi_36b)

```

# combining the two  dataframes above
```{r}
df_ecbi = cbind(df_intensity, df_problem)
```

# Adding data from qualltrics
```{r}

df_qual = read_csv("ECBI_qualtrics.csv")
df2 = df_qual[3:78, 18:92]
df2 = df2 %>% select(-c("Q7"))

df2 = df2 %>% select(2, 1, 3:17, 33:42, 53:63, 18:32, 43:52, 64:74)

initial_names= colnames(df2)

new_names = colnames(df_ecbi)
names(df2) <- new_names

# problem scores to numeric
df2[39:74] = ifelse(df2[39:74] == "Yes", 1, 0)

df_ecbi2 = rbind(df2, df_ecbi)
  
```

# adding calculated columns

## Intensity

```{r}

df_int2 = df_ecbi2 %>% select(1:38)

df_int2[3:38] <- sapply(df_int2[3:38],as.numeric)
sapply(df_int2, class)

# intensity scores

## calculating number missing
df_int2$numMissing_intensity <- rowSums(is.na(df_int2))

## calculating sum
df_int2$intensity_total[df_int2$numMissing_intensity < 3] = rowSums(df_int2[df_int2$numMissing_intensity < 3, c(3:38)], na.rm =  TRUE)
```


## problem scores

```{r}

df_prob2 = df_ecbi2 %>% select(39:74)

## calculating number missing
df_prob2$numMissing_problem <- rowSums(is.na(df_prob2))

## calculating sum
df_prob2$problem_total[df_prob2$numMissing_problem < 3] = rowSums(df_prob2[df_prob2$numMissing_problem < 3, c(1:36)], na.rm =  TRUE)
```

# ordering data frame
```{r}

# merging problem and intensity dfs again
df_ecbi3 = cbind(df_int2, df_prob2)

df_ecbi_ordered <- df_ecbi3[order(df_ecbi3$Child_ID),]

```

```{r}


table(df_ecbi_ordered$Child_ID)

df_ecbi_ordered$Child_ID = recode(df_ecbi_ordered$Child_ID, "bc" = "239", "BLC" = "239", "BT" = "230", "EB" = "231", "eb" = "231", "Eb" = "231", "EJB" = "231", "ep" = "233", "EP" = "233", "gaj" = "234", "GAJ" = "234", "GJ" = "234", "LK"  = "235", "sctm" = "236", "SCTM" = "236", "XJ" = "237", "zlm" = "238", "ZLM" = "238", "zm" = "238", "ZM" = "238")

df_ecbi_ordered = df_ecbi_ordered %>% filter(Child_ID != "test")


### 207 causing problems - was being pooled across wualltrics and REDCap - removing it for now
###df_ecbi_ordered = df_ecbi_ordered %>% filter(Child_ID != "207")
```

# create timepoint column
```{r}

df_ecbi_ordered = df_ecbi_ordered %>%
    group_by(Child_ID) %>%
    mutate(timepoint = 1:n()) %>%
    ungroup

```

# Difference scores for problem

```{r}

df_problem = df_ecbi_ordered %>% select("Child_ID", "problem_total", "timepoint")
df_problem$problem_total_z = as.numeric(scale(df_problem$problem_total, center = TRUE, scale = TRUE))

diff_df <- df_problem  %>% na.omit() %>%           
   group_by(Child_ID) %>%
  mutate(last3mean = mean(tail(problem_total_z, 3), na.rm = TRUE),
         first3mean = mean(head(problem_total_z, 3), na.rm = TRUE),
         change = first3mean - last3mean)

count = diff_df %>% group_by(Child_ID) %>% summarise(n = n()) 

                          
problem_final_df =diff_df %>% group_by(Child_ID) %>% sample_n(1) %>% select(-c(problem_total, timepoint, problem_total_z)) %>% cbind(., count [,2])

write.csv(problem_final_df, "problem_ecbi.csv")

## qualitative graph of scores
ggplot(data = df_problem, aes(x = timepoint, y= problem_total)) +
  geom_point() +
  facet_wrap(facets = "Child_ID")
  
```


## Graph
```{r}

# scores across time
longitudinal = ggplot(df_problem, aes(x = timepoint, y = problem_total_z, color = as.character(Child_ID))) +
  geom_point() +
  geom_line() +
  stat_smooth(method = "lm", se = FALSE)

longitudinal

# number of sessions vs change scores

## LINEAR
graph2 = ggplot(problem_final_df, aes(x = n, y = change)) +
  geom_point(alpha =0.3) +
  geom_line(alpha =0.2) +
  geom_smooth(method = "lm", color = "orange") +
  theme_bw() +
  xlab("Number of sessions") +
  ylab("Problem Change Scores")
graph2


 
summary(lm(data = problem_final_df, change ~ n))

```

# Difference scores for intensity

```{r}

df_intensity = df_ecbi_ordered %>% select("Child_ID", "intensity_total", "timepoint")
df_intensity$intensity_total_z = as.numeric(scale(df_intensity$intensity_total, center = TRUE, scale = TRUE))


diff_df_intensity <- df_intensity  %>% na.omit() %>%           
   group_by(Child_ID) %>%
  mutate(last3mean = mean(tail(intensity_total_z, 3), na.rm = TRUE),
         first3mean = mean(head(intensity_total_z, 3), na.rm = TRUE),
         change = first3mean - last3mean)

count_intensity = diff_df_intensity %>% group_by(Child_ID) %>% summarise(n = n()) 

                          
intensity_final_df =diff_df_intensity %>% group_by(Child_ID) %>% sample_n(1) %>% select(-c(intensity_total, timepoint, intensity_total_z)) %>% cbind(., count_intensity [,2])

write.csv(intensity_final_df, "intensity_ecbi.csv")

## qualitative graph of scores
ggplot(data = df_intensity, aes(x = timepoint, y= intensity_total)) +
  geom_point() +
  facet_wrap(facets = "Child_ID")
  
  
```

## Graphs
```{r}

# scores across time
longitudinal_intensity = ggplot(df_intensity, aes(x = timepoint, y = intensity_total_z, color = as.character(Child_ID))) +
  geom_point() +
  geom_line() +
  stat_smooth(method = "lm", se = FALSE)

longitudinal_intensity

# number of sessions vs change scores
graph2_intensity = ggplot(intensity_final_df, aes(x = n, y = change)) +
  geom_point(alpha = 0.3, color = "purple") +
  geom_line(alpha = 0.2, color = "purple") +
  geom_smooth(method = "lm", color = "purple") +
  geom_point(data = problem_final_df, aes(x = n, y = change), alpha = 0.3, color = "orange") +
  geom_line(data = problem_final_df, aes(x = n, y = change), alpha = 0.2, color = "orange") +
  geom_smooth(data = problem_final_df, aes(x = n, y = change), method = "lm", color = "orange") +
  theme_bw(base_size = 26) +
  xlab("Number of Sessions") +
  ylab("Improvement in Behavior") +
  theme(
   # panel.background = element_rect(fill='transparent'), #transparent panel bg
   plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  #  panel.grid.major = element_blank(), #remove major gridlines
   # panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  ) +
  theme(text=element_text(color="white"),axis.text=element_text(color="white"))
graph2_intensity
ggsave('graph2_intensity.png', graph2_intensity, bg='transparent', width = 10, height = 6)

summary(lm(data = intensity_final_df, change ~ n))

```

# GROWTH CURVE MODELS - PROBLEM SCORES
```{r}

df_problem_clean = na.omit(df_problem)
# Graph
ggplot(df_problem, aes(x = timepoint, y = problem_total_z, group = Child_ID)) +
  geom_line() + # add individual line with transparency
  stat_summary( # add average line
    aes(group = 1),
    fun = mean,
    geom = "line",
    size = 1.5,
    color = "red"
  ) +
  theme_bw() + # nice theme
  labs(x = "Session Number", y = "ECBI Problem Score") # nice labels

#MLM model
library(lme4)
# unconditional means model (a.k.a. random effects model)
m0 <- lmer(data = df_problem, problem_total_z ~ 1 + (1 | Child_ID))
# check results
summary(m0)
# calculating ratio of between and within variation
0.6795/(0.6795+0.2918) #0.669 (66% explained by between person, 34% by within)


# let's look at prediction base on this model
df_problem_clean$pred_m0 <- predict(m0)
df_problem_clean %>% 
  filter(Child_ID %in% 201:205) %>% # select just five individuals
  ggplot(aes(timepoint, pred_m0, color = as.factor(Child_ID))) +
  geom_point(aes(timepoint, problem_total_z)) + # points for observer logincome
  geom_smooth(method = lm, se = FALSE) + # linear line based on prediction
  theme_bw() + # nice theme
  labs(x = "Time point", y = "ECBI Problem Score") + # nice labels
  theme(legend.position = "none") # hide legend


## create new variable starting from 0
df_problem_clean <- mutate(df_problem_clean, timepoint0 = timepoint - 1)
# see how it looks like
head(df_problem_clean, n = 10)

# unconditional change model (a.k.a. MLMC)
m1 <- lmer(data = df_problem_clean, problem_total_z ~ 1 + timepoint0 + (1 | Child_ID))
summary(m1)

# let's look at prediction base on this model
df_problem_clean$pred_m1 <- predict(m1)
df_problem_clean %>% 
  filter(Child_ID %in% 201:205) %>% # select just five individuals
  ggplot(aes(timepoint, pred_m1, color = as.factor(Child_ID))) +
  geom_point(aes(timepoint, problem_total_z)) + # points for observer logincome
  geom_smooth(method = lm, se = FALSE) + # linear line based on prediction
  theme_bw() + # nice theme
  labs(x = "Timepoint", y = "ECBI Problem Score") + # nice labels
  theme(legend.position = "none") # hide legend

# unconditional change model (a.k.a. MLMC) with re for change
m2 <- lmer(data = df_problem_clean, problem_total_z ~ 1 + timepoint0 + (1 + timepoint0 | Child_ID))
summary(m2)

################################################
### explanation
Random effects:
- intercept: inddividual differences in baseline scores (between child difference)
- timepoint: individual differences in change scores (between child differtence)
- residual: within child vaairancce
Fixed effects:
- intercept: average scores at time 0
- wave0: averahe rate of change

################################################
__________________________________________________________________________

df_problem_clean$pred_m2 <- predict(m2)
df_problem_clean %>% 
  filter(Child_ID %in% c(201:206, 208:220, 229:231, 233:238)) %>% # select just two individuals
  ggplot(aes(timepoint, pred_m2, color = as.factor(Child_ID))) +
  geom_point(aes(timepoint, problem_total_z)) + # points for observer logincome
  geom_smooth(method = lm, se = FALSE) + # linear line based on prediction
  theme_bw(base_size = 26) + # nice theme
  labs(x = "Timepoint", y = "ECBI Problem Score") + # nice labels
  theme(legend.position = "none") # hide legend

```


# GROWTH CURVE MODELS - INTENSITY SCORES
```{r}

df_intensity_clean = na.omit(df_intensity)
# Graph
ggplot(df_intensity_clean, aes(x = timepoint, y = intensity_total_z, group = Child_ID)) +
  geom_line() + # add individual line with transparency
  stat_summary( # add average line
    aes(group = 1),
    fun = mean,
    geom = "line",
    size = 1.5,
    color = "red"
  ) +
  theme_bw() + # nice theme
  labs(x = "Session Number", y = "ECBI Intensity Score") # nice labels

# unconditional means model (a.k.a. random effects model)
m0 <- lmer(data = df_intensity_clean, intensity_total_z ~ 1 + (1 | Child_ID))
# check results
summary(m0)
# calculating ratio of between and within variation
0.9317/(0.9317+0.2885) #0.7635 (76% explained by between person, 24% by within)


# let's look at prediction base on this model
df_intensity_clean$pred_m0 <- predict(m0)
df_intensity_clean %>% 
  filter(Child_ID %in% 201:205) %>% # select just five individuals
  ggplot(aes(timepoint, pred_m0, color = as.factor(Child_ID))) +
  geom_point(aes(timepoint, intensity_total_z)) + # points for observer logincome
  geom_smooth(method = lm, se = FALSE) + # linear line based on prediction
  theme_bw() + # nice theme
  labs(x = "Time point", y = "ECBI Intensity Score") + # nice labels
  theme(legend.position = "none") # hide legend


## create new variable starting from 0
df_intensity_clean <- mutate(df_intensity_clean, timepoint0 = timepoint - 1)
# see how it looks like
head(df_intensity_clean, n = 10)

# unconditional change model (a.k.a. MLMC)
m1 <- lmer(data = df_intensity_clean, intensity_total_z ~ 1 + timepoint0 + (1 | Child_ID))
summary(m1)

# let's look at prediction base on this model
df_intensity_clean$pred_m1 <- predict(m1)
df_intensity_clean %>% 
  filter(Child_ID %in% 201:205) %>% # select just five individuals
  ggplot(aes(timepoint, pred_m1, color = as.factor(Child_ID))) +
  geom_point(aes(timepoint, intensity_total_z)) + # points for observer logincome
  geom_smooth(method = lm, se = FALSE) + # linear line based on prediction
  theme_bw() + # nice theme
  labs(x = "Timepoint", y = "ECBI Intensity Score") + # nice labels
  theme(legend.position = "none") # hide legend

# unconditional change model (a.k.a. MLMC) with re for change
m2 <- lmer(data = df_intensity_clean, intensity_total_z ~ 1 + timepoint0 + (1 + timepoint0 | Child_ID))
summary(m2)

df_intensity_clean$pred_m2 <- predict(m2)
df_intensity_clean %>% 
  filter(Child_ID %in% c(201:206, 208:220, 229:231, 233:238)) %>% # select just two individuals
  ggplot(aes(timepoint, pred_m2, color = as.factor(Child_ID))) +
  geom_point(aes(timepoint, intensity_total_z)) + # points for observer logincome
  geom_smooth(method = lm, se = FALSE) + # linear line based on prediction
  theme_bw(base_size = 26) + # nice theme
  labs(x = "Timepoint", y = "ECBI Intensity Score") + # nice labels
  theme(legend.position = "none") # hide legend

```


# Demographics
```{r}

dems = read_csv("demographics.csv")

table(dems$Race)
22/29 #-- white (76%)
5/29 #-- bllack (17%)
1/29 #-- asian (3%)
1/29 #-- dk (3%)

table(dems$SES)
1/29 #<10k (3%)
2/29 #10k-25k (7%)
3/29 #25k-50k (10%)
4/29 #50k-75k (14%)
9/29 #75k-100k (31%) 
7/29 #100k-150k (24%)
1/29 #>150k (3%)
2/29 #dk (7%)

table(dems$gender)
#female - 10 (34%)
10/29
#male - 19 (66%)
19/29

summary(dems$Age)
mean - 4
2-8 range

# Pie chart - gender
mytable <- table(dems$gender)
pie(mytable, labels = NA, col = c('white', 'dark green'))

# Pie chart - race
mytable <- table(dems$Race)
pie(mytable, labels = NA, col = c('white', 'yellow', 'purple', 'blue', 'pink', 'dark green', 'orange'))

# Pie chart - SES
mytable <- table(dems$SES)
pie(mytable, labels = NA,  col = c('white', 'yellow', 'violet', 'blue', 'pink', 'dark green', 'orange'))



# meaan number of sessions
intensity_final_df_207 = intensity_final_df %>% filter(Child_ID != 207)
summary(intensity_final_df$n)
```


# LOOKING AT 207
```{r}

# REDCap data
df_ecbi_test = df_ecbi %>% filter(Child_ID == 207)

# Qualtrics Data
df2_test = df2 %>% filter(Child_ID == 'bc' | Child_ID == 'BLC')

### cannot be same - one in 2019, one in 2022

```



